---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Dennis Raj"
date: "April 26, 2017"
output: html_document
---

```{r Step 1: Import Data and Setup Libraries, include=FALSE }
knitr::opts_chunk$set(echo = TRUE)
#load libraries
library(dplyr)
library(magrittr)
library(readr)
library(ggplot2)
library(ROCR)
library(caTools)


# make sure flights_final.csv is in the current working directory, if not use Rstudio or setwd() 
# set the working directory to read in the csv
flights <- tbl_df(read_csv("flights_final.csv", col_names = TRUE))
```

```{r Step 2: Data Cleaning}

# from the assignment -- predict delays >= 22 minutes

# Filter the data to remove na's and limit the data to arrival delays greater than or equal to 22 minutes 
flights$cat_delay <- "ontime"
flights$cat_delay[flights$arr_delay >= 22] <- "late"
flights$cat_delay <- factor(flights$cat_delay)

#Double check the counts of each of the types of flights
flights %>% select(cat_delay, arr_delay) %>% group_by(cat_delay) %>% 
                            summarize(count = n())




#limit the data to variables of interest

flights_clean <- flights %>% select(cat_delay, month.x, dep_delay, carrier, origin, dest, distance)
```

```{r Step 3: Initial Investigation}

ggplot(data = flights_clean, aes(cat_delay)) + geom_histogram(stat="count")

split <- sample.split(flights_clean$cat_delay, SplitRatio = 0.75)
flights_train <- subset(flights_clean, split==TRUE)
flights_test <- subset(flights_clean, split == FALSE)

#logistic regression model - first test

fit_start <- glm(formula = cat_delay ~ month.x + dep_delay + carrier + origin + dest + distance, data = flights_train , family = binomial(link = "logit"))

#summary of the fit
summary(fit_start)

#testing the anova of the fit using Chi-squared
anova(fit_start, test = 'Chisq')

#plotting the fit
plot(fit_start)

#testing the fit with the new data
fit_test <- glm(formula = cat_delay ~ month.x + dep_delay + carrier + origin + dest + distance, data = flights_test , family = binomial(link = "logit"))

#summary of the fit
summary(fit_test)

#testing the anova of the new fit
anova(fit_test, test = 'Chisq')

#plotting the new fit
plot(fit_test)

```

## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

```


# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   


# Question:

Is this a good model?  (Write your answer here.)
Probably not, since I wasn't able to complete this assignment. 

# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

```{r Step 4: Testing the model with limited data}

flights_update <- flights %>% select(cat_delay, month.x, carrier, origin, dest)

fit_update <- glm(formula = cat_delay ~ month.x + carrier + origin + dest, data = flights_clean, family = binomial(link = "logit"))

summary(fit_update)

anova(fit_update, test = 'Chisq')

plot(fit_update)
```
