---
title: "Classification Metrics"
author: "Dennis Raj"
date: "May 9, 2017"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 

## Model
```{r "model"}
library(dplyr)
library(readr)
library(ggplot2)
flights <- tbl_df(read_csv("flights_clean.csv", col_names = TRUE))
model <- glm(formula = late ~ dep_delay + month.x + air_time + distance + carrier, family = binomial(), data = flights)
flights$pred <- predict(model, flights, type="response")
flights$predict <- "FALSE"
flights$predict[flights$pred >= .5] <- "TRUE" 
  
flights_clean <- flights %>% select(late, predict)
names(flights_clean)[names(flights_clean) == "late"] <- "y"
names(flights_clean)[names(flights_clean) == "predict"] <- "y_hat"  
```

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions"}
library(dplyr)

# true positive rate = TP / TP + FN
tpr <- function(x){
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      tpr_calc <- TP / (TP + FN)
      return(tpr_calc)
}


# False Positive Rate = FP / FP + TN
fpr <- function(x){
      FP <- x %>% filter(y == "FALSE") %>% filter(y_hat == "TRUE") %>% count()
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      fpr_calc <- FP / (FP + TN)
      return(fpr_calc)
}


# True Negative rate = TN / TN + FP
tnr <- function(x){
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      FP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      tnr_calc <- TN / (TN + FP)
      return(tnr_calc)
}

# False Negative rate = FN / TN + FN
fnr <- function(x){
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      fnr_calc <- FN / (TN + FN)
}


# Sensitivity = True Positive Rate
sensitivity <- function(x){
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      sensitivity_calc <- TP / (TP + FN)
      return(sensitivity_calc)
}

# Specificity = True Negative Rate
specificity <- function(x){
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      FP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      specificity_calc <- TN / (TN + FP)
      return(specificity_calc)
}

# Recall = True Positive Rate = TP / TP + FN
f.recall <- function(x){
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      recall_calc <- TP / (TP + FN)
      return(recall_calc)
}

# Prescision = positive predictive value = TP / TP + FP
prescision <- function(x){
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      FP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      prescision_calc <- TP / (TP + FP)
      return(prescision_calc)
  
}

# Prevalence =  observed true / observed true + observed false
prevalence <- function(x){
      OT <- x %>% filter(y == "TRUE") %>% count()
      OF <- x %>% filter(y == "FALSE") %>% count()
      prevalence_calc <- OT / (OT + OF)
      return(prevalence_calc)
}

# Accuracy = TP + TN / TP + TN + FP + FN
accuracy <- function(x){
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      FP <- x %>% filter(y == "FALSE") %>% filter(y_hat == "TRUE") %>% count()
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      accuracy_calc <- (TP + TN) / (TP + TN + FP + FN)
      return(accuracy_calc)
}
test <- accuracy(flights_clean)
# Kappa = (Proportion actual - Proportion expected) / 1 - Proportion expected

f.kappa <- function(x){
      #baselines to calculate rates
      TP <- x %>% filter(y == "TRUE") %>% filter(y_hat == "TRUE") %>% count()
      TN <- x %>% filter(y == "FALSE") %>% filter(y_hat == "FALSE") %>% count()
      FP <- x %>% filter(y == "FALSE") %>% filter(y_hat == "TRUE") %>% count()
      FN <- x %>% filter(y == "TRUE") %>% filter(y_hat == "FALSE") %>% count()
      total_count <- (TP + TN + FP + FN)
      
      #calculating rates
      TP_rate <- TP / total_count
      TN_rate <- TN / total_count
      
      #calculating PR(a)
      pr_a <- TP_rate + TN_rate
      
      #calculating PR(e)
      pred_true <- x %>% filter(y_hat == "TRUE") %>% count()
      pred_false <- x %>% filter(y_hat == "FALSE") %>% count()
      pred_true_rate <- pred_true / total_count
      pred_false_rate <- pred_false / total_count
      actual_true <- x %>% filter(y == "TRUE") %>% count()
      actual_false <- x %>% filter(y == "FALSE") %>% count()
      actual_true_rate <- actual_true / total_count
      actual_false_rate <- actual_false / total_count
      
      pr_e <- (pred_true_rate * actual_true_rate) + (pred_false_rate * actual_false_rate)
      # final k calculation  
      k <- (pr_a - pr_e) / (1 - pr_e)
      return(k)
}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}
# True Positive Rate
tpr(flights_clean)

# False Positive Rate
fpr(flights_clean)


# True Negative Rate
tnr(flights_clean)

# False Negative Rate 
fnr(flights_clean)

# Sensitivity
sensitivity(flights_clean)

# Specificity
specificity(flights_clean)

# Recall
f.recall(flights_clean)

# Prescision
prescision(flights_clean)

# Prevalence
prevalence(flights_clean)

# Accuracy
accuracy(flights_clean)

# Kappa
f.kappa(flights_clean)

```

