---
title: "Caret / Recursive Partitioning"
author: "Dennis Raj"
date: "May 16, 2017"
output: pdf_document
---

```{r init, warning=FALSE, echo=FALSE, message=FALSE}
library(rpart)
library(caret)
library(readr)
library(dplyr)
```


## Exercise 1: caret/logistic regression (5 points)

Rebuild your logistic regression model from the previous week, this time using the `caret` package. 

- Calculate the training or apparent performance of the model. 
- Calculate an unbiased measure of performance 
- Create a ROC Curve for your model

Show all work.

```{r logistic model}
flights <- tbl_df(read_csv("flights_clean.csv", col_names = TRUE))
flights$month.x <- as.factor(flights$month.x)
flights$carrier <- as.factor(flights$carrier)
flights$origin <- as.factor(flights$origin)
flights$dest <- as.factor(flights$dest)
glimpse(flights)

flights_clean <- flights %>% select(month.x, dep_delay, carrier, origin, dest, distance, late)

set.seed(1234)

test <- train(Class ~ ., 
              data = flights_clean, 
              method = "glm",
              verbose = TRUE)

```


## Exercise 2: caret/rpart (5 points)

Using the `caret` and `rpart` packages, create a **classification** model for flight delays using your NYC FLight data. Your solution should include:

- The use of `caret` and `rpart` to train a model.
- An articulation of the the problem your are 
- An naive model
- An unbiased calculation of the performance metric
- A plot of your model -- (the actual tree; there are several ways to do this)
- A discussion of your model 



Show and describe all work

```{r rpart model}
# Your Work Here
```


### Questions:

- Discuss the difference between the models and why you would use one model over the other?
- How might you produce an ROC type curve for the *rpart* model? 
