---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "Dennis Raj"
date: "April 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: Import Data & Setup Libraries

```{r "Step 1: Import Data and Setup Libraries" }
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
# make sure flights_final.csv is in the current working directory, if not use Rstudio or setwd() 
# set the working directory to read in the csv
flights <- tbl_df(read_csv("flights_final.csv", col_names = TRUE))

#double check the structure of flights
glimpse(flights)
```


## Step 2: Visual Exploration
```{r "Step 2: Visual Exploration" }

# visualize the dependent variable (arrival delays)
# bin width set to 60 minutes arbitrarily to represent hours
flights$arr_delay %>% qplot( , binwidth = 60)


#graph arrival delays, discuss learnings from graph

#plot potential predictors with simple scatter plot 
#month, dep_delay, carrier
#origin
#dest
#air_time
#distance
#year of plane
#type of plane
#engine
#temp
#wind_dir
#wind_gust
#precip
#pressure
#visib
#faa for airport name

```

## Step 3: Prep the data 
```{r "Step 3: Prep the data" }
#remove NA

```

## Step 4: Naive Model
```{r "Step 4: Naive Model" }


```

## Step 6: Create a training sample
```{r "Step 5: Train the Model" }


```


## Step 7: Evaluate the model
```{r "Step 5: Train the Model" }


```



# Question:

Is this a good model?  (Write your answer here.)

